Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2018-02-21 paper.tex
Write Test D: test gr terms

\subsection{Idealized Compact Star:
  Testing Gravitational Redshift and Geodesic Curvature}
\label{ssec:test_gravity}
Unfinished: describe and point readers to Asano-Fukuyama test presented
in thesis ~\cite[Sec.~4.3.3]{deat2015-thesis}.
Also compare a few geodesics in Kerr to analytic.
\footnote{TODO (Brett): compute some geodesics in Kerr}

>>>>

\subsection{Idealized Compact Star:
  Testing Gravitational Redshift and Geodesic Curvature}
\label{ssec:test_gravity}
To test the general relativistic terms in our formulation, which account for
the gravitational redshift and geodesic curvature of the neutrinos,
we sample distribution functions outside an idealized hot compact star,
and compute a neutrino-antineutrino interaction integral describing the
energy deposited per time per volume due to the process
$\nu \bar{\nu} \rightarrow e^{-} e^{+}$.
We describe this code test in detail in
\cite[Sec.~4.3]{deat2015-thesis}, and here give a brief summary.

The $\nu \bar{\nu}$-annihilation integral outside a compact star was computed
semi-analytically in \cite{asan2000-nunubar}. Since then many studies of
$\nu \bar{\nu}$-annihilation in more realistic configurations have used the
compact star as a standard code test \cite{birk2007-nunubar,
  hari2010-gr_nunubar_collapsar,zala2011-nunubar}.
We compute the power density (energy per time per volume) due to
$\nu\bar{\nu}$-annihilation measured by a stationary observer above the star
using:
\begin{equation}
  \label{eqn:q_nunubar}
  q_{\nu\bar{\nu}} = A \int d^3p_{\nu} d^3p_{\bar{\nu}}
  f_{\nu}(p_{\nu j})f_{\bar{\nu}}(p_{\bar{\nu} k})
  \frac{p_{\nu t}+p_{\bar{\nu} t}}{p_{\nu}^t p_{\bar{\nu}}^t}
  (p_{\nu\alpha}p_{\bar{\nu}}^\alpha)^2
\end{equation}
where as in \cite{asan2000-nunubar} we account for the energy redshift to
infinite separation by the energy weighting $p_{\nu t}+p_{\bar{\nu} t}$,
$A=2 c^3 K G_{\rm F}^2$,
the Fermi constant is $G_{\rm F}=5.29\times10^{-44}\,{\rm cm}^2\,{\rm MeV}^{-2}$,
\begin{equation}
  K \left\{ {e \atop \mu\,\tau} \right\} = 
  \frac{1}{6\pi}
  \left(1 \, \left\{ {+ \atop -} \right\}
  \, 4 \sin^2 \theta_w + 8 \sin^4 \theta_w\right)
\end{equation}
and the weak mixing angle is $\sin^2\theta_w=0.23$.

Because Eqn.~\ref{eqn:q_nunubar} has such high dimension,
a simple unigrid integral solution---sampling $f_\nu$ and $f_{\bar{\nu}}$
over fixed step sizes in momentum space---is impractical.
We compute the integral using the adaptive Monte Carlo Vegas technique
\cite{pres2007-nr_3rd_ed},
which iteratively samples those regions of momentum space that contribute
most to the integral. At each iterative stage the algorithm estimates the
error, and terminates when some error threshold is achieved.

In order to stress-test the gravity-dependence of the code,
we choose an unphysically compact star configuration with
radius $R=4.43$~km,
in a Schwarzschild metric with
gravitational radius $R_g=2.95$~km.
To compare to the calculation in \cite{asan2000-nunubar},
instead of integrating the formal solution for $f_\nu$ and $f_{\bar{\nu}}$
using Eqns.~\ref{eqn:rendering_fae} and \ref{eqn:rendering_fse},
we compute only the boundary term using Eqn.~\ref{eqn:rendering_fbdry},
and neglect the attenuation due to the optical depth,
This method is equivalent to transporting the neutrino distribution function
in a state of radiative equilibrium with the matter in the star
up to the observer assuming no interactions along the trajectory.
To define the neutrino distribution function in the star,
we make the star homogeneous, with
temperature $T=5$~MeV and
chemical potentials $\eta_{\nu_e}=-\eta_{\bar{\nu}_e}=0.1$,
and we assume stationary fluid.

% Data from
% DATA=$RUNDIR//q.dat
% RUNDIR=$RESEARCHDIR/20150423-thesis_data_and_figures/NeutrinoAnnihilationConvergence/AsanoFukuyamaSphere/Run-01.0/Lev2
% RESEARCHDIR=$HOME/Documents/ResearchGraveyard/NumericalRelativity/DataAndVis
% The integrals computed were the following 12:
% qs={6.87957*^27, 6.83298*^27, 6.76573*^27, 7.02744*^27, 6.88277*^27, 6.89031*^27,
%     6.84114*^27, 6.85656*^27, 6.90977*^27, 6.82969*^27, 6.83426*^27, 6.93181*^27}
The power density deposited by this interaction at a coordinate radius $r=7.38$~km
is computed from formulae in \cite{asan2000-nunubar} as
\begin{equation}
  q_{\nu\bar{\nu}}=6.89\times10^{27}\,{\rm erg}\,{\rm cm^3}\,{\rm s}^{-1}.
\end{equation}
We computed the integral twelve times at an error threshold of 1\% and measured
a mean of
\begin{equation}
  q_{\nu\bar{\nu}}=6.87\pm0.07\times10^{27}\,{\rm erg}\,{\rm cm^3}\,{\rm s}^{-1},
\end{equation}
with the error bars expressing the standard deviation between the twelve
calculations. Each run computed the integral using $N$ samples of the integrand
(requiring $2N$ rays, one for each sample of $f_\nu$ and $f_{\bar{\nu}}$),
with $N$ ranging from 56,000 to 72,000.

The success of this test gives us confidence in the code's ability to handle a
general spacetime metric, since errors in gravitational redshift would have
affected samples of $f$ in the integrand (e.g.\ sampling the distribution function
at the wrong local energy), and errors in geodesic integration would have
affected the angular size of the star (e.g.\ causing the star to look larger
or smaller).
